Loop-Aware Optimizations in PyPy's Tracing JIT

### 2012-08-08 talk/dls2012/paper.tex
add acknowledgements
start expanding the explanation of the benchmarks

\subsection{Python}
The Python interpreter of the RPython framework is a complete Python
version 2.7 compatible interpreter. A set of numerical
calculations were implemented in both Python and in C and their
runtimes are compared in Figure~\ref{fig:benchmarks}. The benchmarks are
\begin{itemize}
\item {\bf sqrt}: approximates the square root of $y$ as $x_\infty$
  with $x_0=y/2$ and $x_k = \left( x_{k-1} + y/x_{k-1} \right) /
  2$. There are three different versions of this benchmark where $x_k$
  is represented with different type of objects: int's, float's and
  Fix16's. The latter, Fix16, is a custom class that implements
  fixpoint arithmetic with 16 bits precision. In Python there is only
  a single implementation of the benchmark that gets specialized
  depending on the class of it's input argument, $y$, while in C,
  there are three different implementations.
\item {\bf conv3}: one-dimensional convolution with fixed kernel-size $3$.
\item {\bf conv5}: one-dimensional convolution with fixed kernel-size $5$.
\item {\bf conv3x3}: two-dimensional convolution with kernel of fixed
  size $3 \times 3$ using a custom class to represent two-dimensional
  arrays.
\item {\bf dilate3x3}: two-dimensional dilation with kernel of fixed
  size $3 \times 3$. This is similar to convolution but instead of
  summing over the elements, the maximum is taken. That places a
  external call to a max function within the loop that prevents some
  of the optimizations.
\item {\bf sobel}: a low-level video processing algorithm used to
  locate edges in an image. It calculates the gradient magnitude
  using sobel derivatives. 
\end{itemize}

@@

%\acks
%Acknowledgments, if needed.

>>>>

\subsection{Python}
The Python interpreter of the RPython framework is a complete Python
version 2.7 compatible interpreter. A set of numerical
calculations were implemented in both Python and in C and their
runtimes are compared in Figure~\ref{fig:benchmarks}. The benchmarks are
\begin{itemize}
\item {\bf sqrt}: approximates the square root of $y$. The approximation is 
initiated to $x_0=y/2$ and the benchmark consists of a single loop updating this
approximation using $x_i = \left( x_{i-1} + y/x_{i-1} \right) / 2$ for $1\leq i < 10^8$. 
Only the latest calculated value $x_i$ is kept alive as a local variable within the loop.
There are three different versions of this benchmark where $x_i$
  is represented with different type of objects: int's, float's and
  Fix16's. The latter, Fix16, is a custom class that implements
  fixpoint arithmetic with 16 bits precision. In Python there is only
  a single implementation of the benchmark that gets specialized
  depending on the class of it's input argument, $y$, while in C,
  there are three different implementations.
\item {\bf conv3}: one-dimensional convolution with fixed kernel-size $3$. A single loop
is used to calculate a vector ${\bf b} = \left(b_1, \cdots, b_n\right)$ from a vector
${\bf a} = \left(a_1, \cdots, a_n\right)$ and a kernel ${\bf k} = \left(k_1, k_2, k_3\right)$ using 
$b_i = k_3 a_i + k_2 a_{i+1} + k_1 a_{i+2}$ for $1 \leq i \leq n$. Both the output vector, $\bf b$, 
and the input vectors, $\bf a$ and $\bf k$, are allocated prior to running the benchmark. It is executed 
with $n=10^5$ and $n=10^6$.
\item {\bf conv5}: one-dimensional convolution with fixed kernel-size $5$. Similar to conv3, but with 
${\bf k} = \left(k_1, k_2, k_3, k_4, k_5\right)$. The enumeration of the elements in $\bf k$ is still 
hardcoded into the implementation making the benchmark consist of a single loop too.
\item {\bf conv3x3}: two-dimensional convolution with kernel of fixed
  size $3 \times 3$ using a custom class to represent two-dimensional
  arrays.
\item {\bf dilate3x3}: two-dimensional dilation with kernel of fixed
  size $3 \times 3$. This is similar to convolution but instead of
  summing over the elements, the maximum is taken. That places a
  external call to a max function within the loop that prevents some
  of the optimizations.
\item {\bf sobel}: a low-level video processing algorithm used to
  locate edges in an image. It calculates the gradient magnitude
  using sobel derivatives. 
\end{itemize}

@@

\acks
We would like to thank Samuele Pedroni, Sven Hager and the anonymous reviewers
for helpful comments on drafts of this paper.
